Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance and associated polyandry

### 2017-09-21 pre_post.tex
Address the statistical inference comment
Finished draft of all changes

\noindent 7226 words (excluding abstract, references, and legends)

@@

Phenotypic values of the three reproductive strategy traits each incur set costs that combine to independently increase the probability that a focal female will die before mating. Numerous forms and mechanisms of direct costs of reproductive strategies could be hypothesised and modelled; the most appropriate formulation depends on the question (see Discussion). By allowing costs of the three strategies to be directly and independently controlled, our model facilitates direct comparison of evolution of pre-copulatory versus post-copulatory inbreeding avoidance given known cost asymmetries. Qualitatively, such costs on female survival probability are biologically reasonable. For example, polyandrous females that undertake increased mate searching or mating can experience increased predation risk \cite[e.g.,][]{Rowe1988, Ronkainen1994, Koga1998}. Females that express pre-copulatory choice can increase the risk of mortality due to harm caused by sexual conflict over mating, and also risk complete mating failure \cite[which equates to pre-reproductive mortality in semelparous organisms;][]{Rowe1994, Kokko2013}. Finally, females that express post-copulatory choice can pay up-front energetic costs, which might result in trade-offs with survival due to developing physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation \cite[e.g.,][]{Gasparini2011, Tuni2013, Fitzpatrick2014a}.

@@

When polyandry alleles ($P_{a}$) were fixed to be positive so that all females were expected to mate multiply and post-copulatory inbreeding allele ($F_{a}$) values were fixed for adaptive inbreeding avoidance, pre-copulatory inbreeding avoidance evolved (i.e., $M_{a}$ values became increasingly negative; Fig. 4A). Such evolution still occurred, but to a much smaller degree, when pre-copulatory inbreeding avoidance was costly (Fig. 4B). However, after $40000$ generations, $M_{a}$ values were less negative when post-copulatory inbreeding avoidance and polyandry were fixed at non-zero values than when they also evolved from initial values of zero ($-15.45$ vs. $-17.62$; compare the blue lines in Fig.s 4A versus Fig. 1B). This implies that pre-existence of fixed post-copulatory inbreeding avoidance can weaken selection and subsequent evolution of pre-copulatory inbreeding avoidance, but does not necessarily preclude it.

>>>>

\noindent \textcolor{red}{\st{7226}} \textcolor{blue}{7267} words (excluding abstract, references, and legends)

@@

Phenotypic values of the three reproductive strategy traits each incur set costs that combine to independently increase the probability that a focal female will die before mating \textcolor{blue}{(realisation of costs precedes mating and fertilisation, so we present further details of mating and fertilisation below).} Numerous forms and mechanisms of direct costs of reproductive strategies could be hypothesised and modelled; the most appropriate formulation depends on the question (see Discussion). By allowing costs of the three strategies to be directly and independently controlled, our model facilitates direct comparison of evolution of pre-copulatory versus post-copulatory inbreeding avoidance given known cost asymmetries. Qualitatively, such costs on female survival probability are biologically reasonable. For example, polyandrous females that undertake increased mate searching or mating can experience increased predation risk \cite[e.g.,][]{Rowe1988, Ronkainen1994, Koga1998}. Females that express pre-copulatory choice can increase the risk of mortality due to harm caused by sexual conflict over mating, and also risk complete mating failure \cite[which equates to pre-reproductive mortality in semelparous organisms;][]{Rowe1994, Kokko2013}. Finally, females that express post-copulatory choice can pay up-front energetic costs, which might result in trade-offs with survival due to developing physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation \cite[e.g.,][]{Gasparini2011, Tuni2013, Fitzpatrick2014a}.

@@

When polyandry alleles ($P_{a}$) were fixed to be positive so that all females were expected to mate multiply and post-copulatory inbreeding allele ($F_{a}$) values were fixed for adaptive inbreeding avoidance, pre-copulatory inbreeding avoidance evolved (i.e., $M_{a}$ values became increasingly negative; Fig. 4A). Such evolution still occurred, but to a much smaller degree, when pre-copulatory inbreeding avoidance was costly (Fig. 4B). However, after $40000$ generations, $M_{a}$ values were less negative when post-copulatory inbreeding avoidance and polyandry were fixed at non-zero values than when they also evolved from initial values of zero ($-15.45$ vs. $-17.62$\textcolor{blue}{, SEs $\approx$ $1.8$ \& $1.6$, respectively}; compare the blue lines in Fig.s 4A versus Fig. 1B). This implies that pre-existence of fixed post-copulatory inbreeding avoidance can weaken selection and subsequent evolution of pre-copulatory inbreeding avoidance, but does not necessarily preclude it.
